---
title: "Readme"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview  
This project is for the analysis of the [UCI Bike Sharind Dataset](https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset), with a focus on generating separate output files in R Markdown.  In particular, we are trying to separately model the daily rider count `cnt` for each day of the week.  

This file contains the script to generate multiple reports.  The report script itself is located [here](Daily Modeling Report.Rmd)

# Reports  

Reports can be found at the links below.  Format is structured identically across reports, with data and statistics corresponding to the day itself.  

[Sunday](Sunday.md)
[Monday](Monday.md)
[Tuesday](Tuesday.md)
[Wednesday](Wednesday.md)
[Thursday](Thursday.md)
[Friday](Friday.md)
[Saturday](Saturday.md)


## Required Packages  


```{r package setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
library(haven)
library(rmarkdown)

options(dplyr.print_min = 5)
options(tibble.print_min = 5)
opts_chunk$set(message = FALSE, cache = TRUE)
```

## 




```{r Parametric Report Generation}
#Create a vector of filenames for outputs
filenames <- c('Sunday.md','Monday.md','Tuesday.md', 
              'Wednesday.md', 'Thursday.md', 'Friday.md', 'Saturday.md')

#For parametric script evaluation, We need individual lists
#Need a labelled DOW field scaling from 0 (Sunday) to 6 (Saturday)
params = lapply(0:6, FUN = function(x){list(DOW = x)})

#Create a tibble so we can iterate along both filenames and parametric values together.
dfDailyReport<- tibble(filenames,params)

#Generate the reports
apply(dfDailyReport, MARGIN=1, FUN = function(x){
  render(input = 'Daily Modeling Report.Rmd', output_file = x[[1]], params = x[[2]])
})



```